---
title: "Lake Water Quality Analysis"
author: "Matthew Ross"
date: "9/17/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties

```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% mdy(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- clarity_only %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')


```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')
```


# Class work

## 3) What is the correlation between Secchi Disk Depth and Chlorophyll a for
sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r}
#Your code here
q3 <- nutr %>%
  group_by(., lagoslakeid) %>%
  summarise(count = n()) %>%
  filter(count >= 200)

nutr200 <- nutr %>%
  filter(lagoslakeid %in% q3$lagoslakeid)

ggplot(nutr200, aes(x = chla, y= secchi)) +
  geom_point()
  
  

```


## Why might this be the case? 

#High chla concentration would result in a like with low clarity becuase algae add particles to the water

## 2) What states have the most data? 
#Minnesota, Wisconsin, and Michigan

### 2a) First you will need to make a lagos spatial dataset that has the total 
number of counts per site.

```{r}

q2 <- nutr %>%
  group_by(., lagoslakeid) %>%
  summarise(count = n()) %>%
lakes_data <- inner_join(spatial_lakes, q2, key = lagoslakeid) %>%
  filter(count >= 200) 
```


### 2b) Second, you will need to join this point dataset to the us_boundaries 
data. 

```{r}
lakes_lots<- st_join(lakes_data, states) %>%
  arrange(count)

mapview(lakes_lots, zcol = 'count')
```


### 2c) Then you will want to group by state and sum all the observations in that
state and arrange that data from most to least toatl observations per state. 

```{r}
q2lakes <- lakes_lots %>%
  group_by(stusps) %>%
  summarise(sum = sum(count)) %>%
  arrange(desc(sum))
```

##3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200 
observations?

```{r}
nut_secchi <- nutr200 %>%
  select(secchi, lagoslakeid) %>%
  group_by(lagoslakeid) %>%
  summarise(av_secchi = mean(secchi, na.rm = TRUE))
nutr_lots <- inner_join(lakes_data, nut_secchi, by="lagoslakeid") 
nutr_lots <- st_as_sf(nutr_lots) %>%
  filter(!is.na(av_secchi))
mapview(nutr_lots, zcol = "av_secchi")


```


